MyPDDL: Tools for efficiently creating PDDL domains and problems

### 2019-07-03 paper.tex
upd

More usable and probably faster results can be achieved by using the
interface between \pddl and Clojure as a distance calculator,
implemented in the tool \mypddldistance. It reads a problem file into
Clojure and extracts all locations, defined in the \texttt{(:init
  ...)} code block. The Euclidean distances between these locations
are then calculated and written back into a new and now extended copy
of the problem file, using the predicate
\texttt{(distance~?o1~?o2~-~object ?n~-~number)}, which specifies the
distance between two objects. Listing \ref{ls:before-mypddl-distance}
and \ref{ls:after-mypddl-distance} show the \texttt{(:init ...)} code
block of a extended version of \emph{Gary's Huge Problem} before and
after using \mypddl-distance, respectively.

@@

\begin{listing}[H]
\begin{minted}[fontsize=\small,bgcolor=dhscodebg,rulecolor=\color{gray!40},frame=lines,framesep=5\fboxrule,framerule=1pt,tabsize=2]{text}
(:init ...
       (location gary 4 2)
       (location pizza 2 3))
\end{minted}
\caption[Excerpt of the extended problem file before using
\mypddl-distance]{\label{ls:before-mypddl-distance}Excerpt of the
  \texttt{(:init~...)} block of the extended file \emph{Gary's Huge
    Problem} before using \mypddl-distance.}
\end{listing}

\begin{listing}[H]
\begin{minted}[fontsize=\small,bgcolor=dhscodebg,rulecolor=\color{gray!40},frame=lines,framesep=5\fboxrule,framerule=1pt,tabsize=2]{text}
(:init ...
       (location gary 4 2)
       (location pizza 2 3)
       (distance gary gary 0.0)
       (distance gary pizza 2.2361)
       (distance pizza gary 2.2361)
       (distance pizza pizza 0.0))
\end{minted}
\caption[Excerpt of the extended problem file after using
\mypddl-distance]{\label{ls:after-mypddl-distance}After the
  application of \mypddl-distance, the calculated distances are
  inserted in the \texttt{(:init ...)} code block in a copy of the
  problem file.}
\end{listing}

@@

\begin{figure}[h]
\captionsetup{singlelinecheck=off}
  \centering
  \hspace{1.7cm}
  \includegraphics[width=0.85\textwidth]{task_time}
  \caption[Diagram of the task completion time]{Task completion time
    for the type hierarchy tasks. The bars display the geometric mean
    averaged over all participants; each cross~($\times$) represents
    the data value of one participant. The percent values at the
    bottom of the bars show the percentage of users that completed the
    task successfully. The question can be found in the Appendix
    (\ref{splisus}, \ref{store}).}\label{fig:task-completions-agg}
\end{figure}

@@

>>>>

  More usable and probably faster results can be achieved by using the
  interface between \pddl and Clojure as a distance calculator,
  implemented in the tool \mypddldistance. It reads a problem file
  into Clojure and extracts all locations, defined in the
  \texttt{(:init ...)} code block. The Euclidean distances between
  these locations are then calculated and written back into a new and
  now extended copy of the problem file, using the predicate
  \texttt{(distance~?o1~?o2~-~object ?n~-~number)}, which specifies
  the distance between two objects. The code blocks below show the
  \texttt{(:init ...)} block of a \pddl problem file before and after using
  \mypddl-distance.
\begin{minted}[fontsize=\small,bgcolor=dhscodebg,rulecolor=\color{gray!40},frame=lines,framesep=5\fboxrule,framerule=1pt,tabsize=2]{text}
(:init ...
       (location gary 4 2)
       (location pizza 2 3))
\end{minted}
\begin{minted}[fontsize=\small,bgcolor=dhscodebg,rulecolor=\color{gray!40},frame=lines,framesep=5\fboxrule,framerule=1pt,tabsize=2]{text}
(:init ...
       (location gary 4 2)
       (location pizza 2 3)
       (distance gary gary 0.0)
       (distance gary pizza 2.2361)
       (distance pizza gary 2.2361)
       (distance pizza pizza 0.0))
\end{minted}

@@

@@

\begin{figure}[h]
\captionsetup{singlelinecheck=off}
  \centering
  \hspace{1.7cm}
  \includegraphics[width=0.85\textwidth]{task_time}
  \caption[Diagram of the task completion time]{Task completion time
    for the type hierarchy tasks. The bars display the geometric mean
    averaged over all participants; each cross~($\times$) represents
    the data value of one participant. The percent values at the
    bottom of the bars show the percentage of users that completed the
    task successfully. The questions can be found in the
    Appendix~\ref{splisus} and
    \ref{store}.}\label{fig:task-completions-agg}
\end{figure}

@@

\appendix

\section{Tasks}
\label{sub:tasks}
\subsection{Deliberately erroneous Logistics Domain}
\label{logistics}
\begin{minted}[fontsize=\tiny,bgcolor=dhscodebg,rulecolor=\color{gray!40},frame=lines,framesep=5\fboxrule,framerule=1pt,tabsize=2]{text}
;;;; Logistics domain

(define (domain ?logistics)

  (:requirements
    :types)

  (:typing truck airplane motorboat - vehicle
           package vehicle suitcase furniture - thing
           airport garage station - location
           car1 car 2 car3 - vehicle
           city location thing - object)

  (:predicates (in-city ?l - location ?c - city)
               (at ?obj - thing ?l - location)
               (key ?v - vehicle) = true
               (full ?v - vehicle)
               (in ?p - package ??veh - vehicle))

  (:action drive
    :parameters (?t - truck ?from ?to - location ?c - city)
    :precondition (and (at ?tr ?from)
                       (in-city ?from ?c)
                       (incity ?to ?c))
    :effect (and (not (at ?t ?from))
                 (at ?t ?to)))

  (:action fly
    :parameters (?a - airplane ?from ?to - airport)
    :precondition (at ?a ?from)
    :effect (and (n0t (at ?a ?from))
                 (at ?a ?to)))

  (:action fuel
    :parameters (?v - vehicle ?c - city ?to airport)
    :precondition (and (not (full ?v))
                       (in-city ?to ?c)
                       (at ?v ?to))
    :effect (full ?v))

  (:action load
    parameters: (?v - vehicle ?p - package ?l - location)
    precondition: (and (?v ?l)
                       (at ?p ?l))
    :effect (and (ay ?p ?l)
                 (in ?p ?v)))

  (:action unload
    :parameters (?v - vehicle p - package ?l - location)
    :precondition (and (at ?v ?l)
                           ?p ?v)
    :effects (and (not (in ?p ?v))
                  (at ?p - ?l))))
                \end{minted}
The original file can be downloaded at
\url{http://ipc.informatik.uni-freiburg.de/PddlExtension}:

\subsection{Deliberately erroneous Coffee Domain}
\label{coffee}
\begin{minted}[fontsize=\scriptsize,bgcolor=dhscodebg,rulecolor=\color{gray!40},frame=lines,framesep=5\fboxrule,framerule=1pt,tabsize=2]{text}
(define COFFEE

  (requirements
    :typing)

  (:types room - location
                 robot human _ agent
                 furniture door - (at ?l - location)
                 kettle ?coffee cup water - movable
                 location agent movable - object)

  (:predicates (at ?l - location ??o - object)
               (have ?m - movable ?a - agent)
               (hot ?m - movable) = true
               (on ?f - furniture ?m - movable))

  (:action boil
    :parameters (?m - movable \$k - kettle ?a - agent)
    :preconditions (have ?m ?a)
    :effect (hot ?m))

  (:action grip-some
    :parameters (?m - movable ?r - robot ?f - _furniture ?l - location)
    :precondition (and (at ?l ?r)
                       (on ?fu ?m)
                       (at ?l ?f))
    :effect (and (have ?m ?r)))

  (:action move
    :parameters: (?m - movable ?a - agent ?from ?to - location)
    :precondition (or (änd (at ?from ?a)
                           (at ?from ?m))
                           (and (at ?from ?m)
                                (location ?from ?a)))
    :effect (and (not (at ?from ?m))
                 (at ?to ?m)))

  (:action change-room
    :parameters (?from-r ?to-r - room ?a - agent)
    :precondition (at ?fromr ?a) 
    :effect (and (not (at ?from-r ?a))
                 (at ?tor ?a)))

  (:action prep-coffee
    :parameters (?a - agent ?c - cjp ?w - water ?cof - coffee)
    :precondition (and (have ?c ?a)
                       (hot ?w))
    :effect (have ?cof ?a))

  (:action ?hand-over
    :parameters (?m - movable ?a1 - agent ?a2 - agent)
    :precondition (have ?m ?a1))
    :effect (and (not (have ?m ?a1))
                 (have ?m ?a2))))
\end{minted}

\subsection{Planet Splisus}
\label{splisus}
\begin{minted}[fontsize=\scriptsize,bgcolor=dhscodebg,rulecolor=\color{gray!40},frame=lines,framesep=5\fboxrule,framerule=1pt,tabsize=2]{text}
(define (domain splisus)

  (:requirements :typing)

  (:types splis - gid
          spleus - splos
          schprok schlok - splus
          rud mekle - lech
          hulpf hurpf - hupf
          sipsi flipsi hupf - splis
          schmok schkok - splus
          gid splos splus - ruffisplisus
          merle - hupf
          ruffisplisus mak lech - object)

  (:predicates (father-of ?r1 - ruffisplisus ?r2 - ruffisplisus)
               (married ?s1 - splos ?s2 - splis)
               (has-weapon ?h - sipsi)
               (dead ?r1 - ruffisplisus)
               (at ?l - lech ?r - ruffisplisus))

  (:action kill
    :parameters (?l - lech ?r1 - ruffisplisus ?s - splis)
    :precondition (and (at ?l ?r1)
                       (at ?l ?s)
                       (married ?r1 ?s)
                       (has-weapon ?s))
    :effect (and (dead ?r1)
            (not (married ?r1 ?s)))))
\end{minted}

Please answer the following five questions on the society and
structure of Planet Splisus:
\begin{enumerate}
\item Are all \emph{Flipsis} also of the type \emph{Ruffisplisus}?
\item Are all \emph{Merles} also \emph{Splus}?
\item Can a \emph{Spleus} be married to a \emph{Schlok}?
\item Only theoretically: Could a \emph{Hurpf} murder a \emph{Spleus}?
\item Let us assume there are three categories of object types on
  Splisus: places, beings and food. Match the three object types
  \emph{Ruffisplisus}, \emph{Mak} and \emph{Lech} with these
  categories.
\end{enumerate}

\subsection{Store}
\label{store}
\begin{minted}[fontsize=\scriptsize,bgcolor=dhscodebg,rulecolor=\color{gray!40},frame=lines,framesep=5\fboxrule,framerule=1pt,tabsize=2]{text}
(define (domain store)

  (:requirements :typing)

  (:types lala lila - zahls
          blisis blusis - ultri
          iltre lula - nulls
          zahls schwinds - knozi
          minis - lala
          ultri sopple schmitzl - lila
          ultres raglos wexis - lola
          kosta - nulls
          nulls spax - minis
          lola - zahls
          knozi schmus - object)

  (:predicates (product ?k - knozi) ; Produkt
               (workplace ?l1 - lola ?l2 - lala) 
               (product-at ?l1 - lola ?l2 - lila) 
               (cashier ?k - knozi) 
               (customer ?s - spax)
               (owns ?l - lila ?s - spax))

  (:action sell
    :parameters (?p - lila ?z - zahls ?l - lola ?w - wexis ?s - spax)
    :precondition (and (product ?p)
                       (cashier ?z)
                       (product-at ?l ?p)
                       (customer ?s))
    :effect (and (product-at ?w ?p)
                 (not (product-at ?l ?p))
                 (owns ?p ?s))))
\end{minted}
Please answer the following five questions concerning the environment
store:
\begin{enumerate}
\item Are objects of the type \emph{Lula} also of the type
    \emph{Minis}?
\item Are \emph{Spax} and \emph{Schmus} \emph{Zahls}?
\item Is it possible for an Iltre to work at a
    \emph{workplace} of the type \emph{Knozi}?
\item Only theoretically: Could a \emph{Lala} sell a
    \emph{Schmitzl} to a \emph{Kosta}?
\item Let us assume our domain \emph{store} models a
    grocery store. There are three categories: humans, products, and
    places. Can you match these world terms with the object types
    \emph{lila}, \emph{lala}, and \emph{lola} from the domain?
\end{enumerate}
